---
title: "Assignment1"
output: html_document
---

# Q1: Loading and preprocessing the data

```{r echo=TRUE}
suppressMessages(library(lubridate))
suppressMessages(library(tidyr))
suppressMessages(library(dplyr))
suppressMessages(library(ggplot2))
options(scipen=999)

# Load the data
dat1<- read.csv("activity.csv")

# Process/transform the data
dat1$date <- ymd(dat1$date)
str(dat1)
```

# Q2: What is mean total number of steps taken per day?
```{r}
#Calculate the total number of steps taken per day
dat1_sum_by_day<- dat1 %>% 
    group_by(date) %>% 
    drop_na() %>% 
    summarise(sum= sum(steps, na.rm=TRUE))

# Make a histogram of the total number of steps taken each day
hist(dat1_sum_by_day$sum, main = "Histogram of total number of steps taken per day", xlab = "Total number of steps taken per day")

# Calculate and report the mean and median of the total number of steps taken per day
mean <- round(mean(dat1_sum_by_day$sum, na.rm= T))
median <- median(dat1_sum_by_day$sum)
```
Mean total number of steps taken per day is `r mean`.  
Median total number of steps taken per day is `r median`.


# Q3: What is the average daily activity pattern?

```{r}
# Prepare the data
dat1_daily_pattern <- dat1 %>% 
    group_by(interval) %>% 
    drop_na() %>% 
    summarise(mean=mean(steps))

# Make a time series plot
dat1_daily_pattern %>% ggplot(aes(y=mean, x= interval)) +
    geom_line()+
    scale_x_continuous(breaks = seq(0,2355,100))

# Find the maximum 5-minute interval
max<- dat1_daily_pattern[which(dat1_daily_pattern$mean==max(dat1_daily_pattern$mean)),1]

```
The `r max`th 5-minute interval contains the maximum number of steps.  

# Q4: Imputing missing values
```{r}
# Calculate and report the total number of missing values in the dataset
apply(is.na(dat1), 2,sum )

dat1_imputed <- dat1

# Filling in all of the missing values in the dataset
for ( i in 1:nrow(dat1_imputed)) {
    if (is.na(dat1_imputed$steps[i])){
        which_row <- which(dat1_daily_pattern$interval==dat1_imputed$interval[i])
        dat1_imputed$steps[i] <- as.numeric(dat1_daily_pattern[which_row,"mean"])
    }
}
apply(is.na(dat1_imputed), 2,sum )

# Make a histogram of the total number of steps taken each day 
dat1_imputed_sum_by_day<- dat1 %>% 
    group_by(date) %>% 
    drop_na() %>% 
    summarise(sum= sum(steps, na.rm=TRUE))

hist(dat1_imputed_sum_by_day$sum, main = "Histogram of imputed total number of steps taken per day", xlab = "Imputed total number of steps taken per day")

mean_imputed <- round(mean(dat1_imputed_sum_by_day$sum),0)
median_imputed <- median(dat1_imputed_sum_by_day$sum)

```
The total number of missing values in the dataset is 2304.
Mean total number of steps taken per day is `r mean_imputed`.  
Median total number of steps taken per day is `r median_imputed`.


# Q5: Are there differences in activity patterns between weekdays and weekends?
```{r}
# Create a new factor variable in the dataset with two levels
dat1_imputed<- dat1_imputed %>% 
    mutate(weekday=weekdays(date,abbreviate=TRUE)) %>%
    mutate(weekday_factor=case_when(
        weekday %in% c("Sun","Sat") ~ "weekend",
        TRUE ~ "weekday")
    )

# Make a panel plot containing a time series plot
dat1_imputed_daily_pattern <- dat1_imputed %>% 
    group_by(interval,weekday_factor) %>% 
    summarise(mean=mean(steps))

dat1_imputed_daily_pattern %>% ggplot(aes(y=mean, x= interval)) +
    geom_line(aes(col=weekday_factor))+
    facet_grid(weekday_factor~.)+
    scale_x_continuous(breaks = seq(0,2355,100))+
    theme_bw()
    

```

